Understanding Structure and Selectivity in Nanostructured Lyotropic Liquid Crystal Membranes using Molecular Simulations

### 2020-02-27 Ben_Manuscripts/thesis/intro.tex
Beefing up intro

  % TFC membranes review -- "Composite reverse osmosis and nanofiltration membranes"
  \noindent Membrane design is completely dependent on the target particle separation. 
  \begin{itemize}
    \item They can be classified based on the size of the particle they reject.
    \item Microfiltration membranes contain 100-10000 nm pores. They can separate large
    particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
    \item Ultrafiltration membranes have pore sizes about 5-500 nm and are useful for the 
    separation of sugars, proteins, viruses and colloidal materials.~\cite{davey_ultrafiltration_2009}
    \item Nanofiltration have pores on the order of 1 nm in size and can be used for the separation
    of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
    \item Reverse osmosis membranes are dense amorphous polymers with no explicit pores. Their
    dense polymer architecture reject all solutes except water making them useful for 
    separating hydrated ions from water.~\cite{warsinger_review_2018}
    \item We have summarized the uses of different classes of membrane 
    technologies in Figure~\ref{fig:size_regimes}.
    \item To deal with complex aqueous streams, and to prevent excessive membrane fouling, 
    these technologies are often used in series, removing larger particles first.~\cite{goosen_fouling_2005}
  \end{itemize}

@@

  \noindent In amorphous membranes, water flux is modeled using the solution-diffusion 
  model.
  \begin{itemize}
    \item It is hypothesized that water and solute molecules partition into the
    membrane and diffuse across the polymer matrix due to a chemical potential
    gradient and then desorb into the permeate.
    \item The solubility and diffusivity can be described in terms of the 
    diffusive water and solute permeabilities ($P_w$ and $P_s$) respectively
    \item \begin{equation}
    A = \frac{P_wV_w}{\delta_mR_gT}
    \end{equation}
    where $V_w$ is molar volume of water, $R_g$ is the gas constant and $T$ is
    absolute temperature.
    \item Solute flux is modeled as:
    \begin{equation}
    J_s = \frac{P_s}{\delta_m}\Delta c_m = B\Delta c_m
    \end{equation}
    where B is a solute permeability coefficient and \Delta c_m is the concentration
    difference across the selective membrane layer.
    \item Under these assumptions, rejection can be reformulated as
    \begin{equation}
    R = \frac{A}{A + B}
    \end{equation}
    which approaches unity as the solute flux tends towards zero.
  \end{itemize}

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi et al. calculated that tripling membrane permeability
    could reduce energy consumption up to 15\% in a seawater reverse osmosis plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which requires chemical and capital costs for
    post-treatment
    \item Addressing these problems with highly selective separations may lead to 
    more gains.
  \end{itemize}

@@

  Selective separations are useful for a wide range of applications.

  % Put this with the desalination section below maybe
  % morton_environmental_1996,bhojwani_technology_2019
  \textit{Desalination}:
  Creating potable water from seawater or brackish water is of paramount importance
  in water-scarce areas. 
  \begin{itemize}
  	\item Compared with thermal distillation techniques, reverse osmosis has been 
  	shown to be more environmently friendly and economical due to its lower energy
  	requirements.
  	\item Thermal techniques are still preferred where excess waste heat or cheap 
  	thermal energy is available, such as cogeneration plants.
  	\item RO has smaller footprint?
  \end{itemize}

  \textit{Organic Micropollutants}:
  Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, which have adverse effects on human health even at low 
  concentrations\cite{schwarzenbach_challenge_2006}
  \begin{itemize}
    \item Micropollutants include pharmaceutical and personal care products, 
    hormones, pesticides and industrial chemicals which find their way into
    our drinking water supply
    \item Is there any infrastructure built to combat this?
  \end{itemize}

  \textit{Recovery of Valuable Dissolved Species}: We can use highly selective
  membranes in order to recover potentially valueable dissolved species from 
  complex waste streams.

@@

  Amorphous thin film composite (TFC) RO membranes are the current industry standard for
  high purity separations.
  \begin{itemize}  
    \item Mostly desalination (?)
    \item Typically you have a porous support layer for mechanical strength and 
    very thin active layer where the separations occur.
    \item Usually only let water through. Usually have to remineralize the
    water to make it drinkable.
  \end{itemize}

  Operation requires high feed pressures.

  \subsection{Standard Nanofiltration}

@@

  Explicitly defined pores allows much lower feed pressures. However, polydispersity in
  pore size lowers their selectivity relative to RO.

@@

  %BJC: format for following paragraphs maybe: Introduce tech, say how it would work, why it doesn't work 
  Ultrathin-film
  graphene and graphene oxide membranes are an active area of research because
  they are atomically thick and therefore offer potential for extremely high
  permeability membranes.~\cite{humplik_nanostructured_2011} 
  \begin{itemize}  
    \item However, scalable synthesis without introducing microscopic, performance
    degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}
  \end{itemize}

@@

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    \item In theory, this topology is ideal for high flux and highly selective separations.
  \end{itemize}

  Unfortunately, it is a difficult task to align the self-assembled hexagonal
  mesophases into continuous pores that traverse the width of the membrane.
  \begin{itemize}
    \item Although the membranes have shown high experimental selectivity, their 
    flux has been very low due to misalignment of the hexagonal mesophases.
  \end{itemize}

%  Recently, Feng et al. has come up with two methods which can be used to 
%  macroscopically align the hexagonal mesophases. 
  Considerable recent efforts have made progress towards the macroscopic
  alignment of the hexagonal mesophases.
  \begin{itemize}
    \item Feng et al. leverage the magnetic anisotropy of the columnar \
    mesophases in order to control their alignment with a magnetic field~\cite{feng_scalable_2014}
    \item Feng et al. also used an approach caused soft confinement which
    takes advantage of the hexagonal columns preference towards anchoring
    perpendicular to either a PDMS or glass substrate.~\cite{feng_thin_2016}
    \item Finally, Feng et al. designed a third technique which uses a structure
    directing molecule in order to template the assembly of a fatty acid 
    molecule into ordered columnar phases.~\cite{feng_highly_2017}
  \end{itemize}

@@

  The space group of the Q\textsubscript{I} phase configuration formed by
  monomer 1 is uncertain and thought to be either Pn3m or Ia3d.
  \begin{itemize}
    \item 6 Q phase architectures have been identified in small molecule
    amphiphile systems.~\cite{mariani_cubic_1988}
    \item Of these, only 4 are consistent with diffraction data generated
    by the gemini surfactant used here.~\cite{pindzola_cross-linked_2003}
    \item Q$^{230}$ (Ia3d), Q$^{224}$ (Pn3m), Q$^{229}$ (Im3m) or Q$^{212}$ (P4\textsubscript{3}32) phase of type I configuration
    \item The presence of $1 / \sqrt{6}$ and $1 / \sqrt{8}$ peaks rules out the
    Q$^{227}$ (Fd3m) and Q$^{223}$ (Pm3n) configurations.
    \item The Pn3m and Ia3d architectures are the most common~\cite{mariani_cubic_1988,wiesenauer_nanoporous_2012}
    \item The most likely phase is the type I Q$^{230}$ (Ia3d) because it is the
    most common phase observed between lamellar and hexagonal phases and the
    monomeric alkyltrimethylammonium salts used to synthesize the gemini LCs
    exhibit clear Ia3d symmetry.
  \end{itemize}

@@

  There are few molecular simulation studies which study structure and transport
  in LLC systems.
  \begin{itemize}
    \item Mondal et al. used MD simulations to study the self assembly of gemini 
    surfactants and observed the formation of H\textsubscript{II} and Q\textsubscript{I}
    phases depending on water content.\cite{mondal_self-assembly_2013}
    \item Mantha and Yethiraj as well as Roy et al. studied the dynamics of confined 
    water in these systems and showed orders of magnitude differences in their motion
    dependent on system geometry.~\cite{mantha_dynamics_2016,roy_water_2016}
    \item Jackson et al. as well as Manthan et al. have combined experiment with simulation
    in order to show how the choice of monomer head group counterion regulate water 
    dynamics.~\cite{jackson_ion-specific_2018,mantha_counterion-regulated_2018}
    \item Sakamoto et al. and Nada et al. observed the dynamics of water molecules
    and ions and hypothesized transport mechanisms in a simplified model of an LLC
    nanopore.~\cite{sakamoto_development_2018,nada_transport_2020}
  \end{itemize}

>>>>

  % TFC membranes review -- "Composite reverse osmosis and nanofiltration membranes"
  \noindent Membrane design is completely dependent on the target particle separation. 
  \begin{itemize}
    \item Membranes are classified based on the size of the particle they reject.
    \item Microfiltration membranes contain pores with diameters ranging from 100-10000 nm.
    They can separate large particles like bacteria and protozoa.~\cite{ma_functionalized_2014}
    \item Ultrafiltration membranes have pore diameters of about 5-500 nm and are useful for the 
    separation of sugars, proteins, viruses and colloidal materials.~\cite{yanful_ultrafiltration_2009}
    \item Nanofiltration have pores on the order of 1 nm in diameter and can be used for the separation
    of small organic molecules and charged species.~\cite{van_der_bruggen_review_2003}
    \item Reverse osmosis membranes are dense amorphous polymers with no explicit pores. Their
    dense polymer architecture rejects all solutes to a high degree except for water making
    them useful for separating hydrated ions from water.~\cite{warsinger_review_2018}
    \item We have summarized the uses of different classes of membrane 
    technologies in Figure~\ref{fig:size_regimes}.
    \item To deal with complex aqueous streams, and to prevent excessive membrane fouling, 
    these technologies are often used in series, removing larger particles first.~\cite{goosen_fouling_2005}
  \end{itemize}

@@

  In amorphous membranes, water and solute flux are modeled using the solution-diffusion 
  model.
  \begin{itemize}
    \item It is hypothesized that water and solute molecules partition into the
    membrane and diffuse across the polymer matrix due to a chemical potential
    gradient and then desorb into the permeate.
    \item The product of a component's solubility and diffusivity is commonly
    referred to as its permeability, $P$.~\cite{wijmans_solution-diffusion_1995}
    \item $P$ is an intrinsic material property and is thus independent of membrane
    thickness.  % BJC: is intensive a more appropriate word than intrinsic?
    \item The water permeability coefficient \textit{is} dependent on thickness, $\delta_m$,
    and can be formulated in terms of the pure water permeability, $P_w$.
    \item \begin{equation}
    A = \frac{P_wV_w}{\delta_mR_gT}
    \end{equation}
    where $V_w$ is molar volume of water, $R_g$ is the gas constant and $T$ is
    absolute temperature.
    \item Solute flux is modeled as:
    \begin{equation}
    J_s = \frac{P_s}{\delta_m}\Delta c_m = B\Delta c_m
    \end{equation}
    where B is a solute permeability coefficient and $\Delta c_m$ is the concentration
    difference across the selective membrane layer.
    \item Under these assumptions, rejection can be reformulated as
    \begin{equation}
    R = \frac{A}{A + B}
    \end{equation}
    which approaches unity as the solute flux tends towards zero.
  \end{itemize}

  Although there has been considerable focus on creating membranes with high
  permeabilities, there is a case to be made that shifting the focus to membrane
  selectivity may offer a more effective route towards lowering the cost of 
  membrane separations.
  \begin{itemize}
    \item Increasing permeabilities beyond those already achieved may have 
    only a small effect on energy requirements and capital costs. 
    \item Cohen-Tanugi et al. calculated that tripling membrane permeability
    could reduce energy consumption up to 15\% in a seawater reverse osmosis plant.~\cite{cohen-tanugi_quantifying_2014}
    \item Perhaps the biggest detriment to energy consumption is the practical
	need for multiple membrane passes in order to achieve a desired purity.
    \item Additionally, many membranes are not suited for high purity separations
    of small neutral species which necessitates chemical and capital costs for
    post-treatment
    \item Addressing these problems with highly selective separations may lead to 
    more gains.
  \end{itemize}

@@

  Selective separations are useful in a wide range of applications.

  \textit{Desalination}:
  Creating potable water from seawater or brackish water is of paramount importance
  in water-scarce areas. 
  \begin{itemize}
  	\item Compared with thermal distillation techniques, reverse osmosis has been 
  	shown to be more environmently friendly and economical due to its lower energy
  	requirements.\cite{morton_environmental_1996}
  	\item Thermal techniques are still preferred where excess waste heat or cheap 
  	thermal energy is available, such as cogeneration plants.\cite{bhojwani_technology_2019}
  	\item RO has smaller footprint % BJC: need citation
  \end{itemize}

  \textit{Organic Micropollutants}:
  Municipal and industrial wastewaters are contaminated by harmful
  micropollutants, which have adverse effects on human health even at low 
  concentrations\cite{schwarzenbach_challenge_2006}
  \begin{itemize}
    \item Micropollutants include pharmaceuticals, personal care products, 
    hormones, pesticides and industrial chemicals which find their way into
    our drinking water supply
    \item Is there any infrastructure built to combat this?
  \end{itemize}

  \textit{Recovery of Valuable Dissolved Species}: We can use highly selective
  membranes in order to recover potentially valuable dissolved species from 
  complex waste streams.

@@

  Amorphous thin film composite (TFC) RO membranes are the current industry standard for
  high purity separations.
  \begin{itemize}  
    \item Desalination, concentration
    \item Typically you have a porous support layer for mechanical strength and 
    very thin active layer where the separations occur.
    \item Usually only let water through. Usually have to remineralize the
    water to make it drinkable.
  \end{itemize}

  Operation requires high feed pressures. (5 - 120 bar) ~\cite{van_der_bruggen_review_2003}

  \subsection{Standard Nanofiltration Membranes}

@@

  Explicitly defined pores allow lower feed pressures (3 - 20 bar),~\cite{van_der_bruggen_review_2003}
  however, polydispersity in pore size lowers their selectivity relative to RO.

@@

  %BJC: More detail on these sections to come.
  % Probable format for following paragraphs: Introduce tech, say how it would work, then why it doesn't work 
  Ultrathin-film
  graphene and graphene oxide membranes are an active area of research because
  they are atomically thick and therefore offer potential for extremely high
  permeability membranes.~\cite{humplik_nanostructured_2011} 
  \begin{itemize}  
    \item However, scalable synthesis without introducing microscopic, performance
    degrading defects has not yet been achieved.~\cite{cohen-tanugi_multilayer_2016,wei_multilayered_2018}
  \end{itemize}

@@

  The H\textsubscript{II} phase is characterized by hexagonally packed, uniform-sized
  and straight pores.
  \begin{itemize}
    \item The hydrophilic head groups aggregate in the pore centers to create
    cylindrical aqueous channels. 
    \item These pore channels are lined with the chemical functionality of the LLC monomers
    and have the potential to interact with solutes in a chemically-dependent manner.
    \item In theory, this pore topology and geometry is ideal for high flux and 
    highly selective separations.
  \end{itemize}

  Unfortunately, it is a difficult task to align the self-assembled hexagonal
  mesophases into continuous pores that traverse the thickness of the membrane.
  \begin{itemize}
    \item Although the membranes have shown high experimental selectivity, their 
    flux has been very low due to misalignment of the hexagonal mesophases.
  \end{itemize}

%  Recently, Feng et al. has come up with two methods which can be used to 
%  macroscopically align the hexagonal mesophases. 
  Considerable recent efforts have made progress towards the macroscopic
  alignment of the hexagonal mesophases.
  \begin{itemize}
    \item Feng et al. leverage the magnetic anisotropy of the columnar \
    mesophases in order to control their alignment with a magnetic field.~\cite{feng_scalable_2014}
    \item Feng et al. also used an approach called soft confinement which
    takes advantage of the hexagonal columns preference towards anchoring
    perpendicular to either a PDMS or glass substrate.~\cite{feng_thin_2016}
    \item Finally, Feng et al. designed a third technique which uses a structure
    directing molecule in order to template the assembly of a fatty acid 
    molecule into ordered columnar phases.~\cite{feng_highly_2017}
  \end{itemize}

@@

  The space group of the Q\textsubscript{I} phase configuration formed by
  monomer 1 is uncertain and thought to be either Pn3m or Ia3d.  %BJC: will have accompanying figure
  \begin{itemize}
    \item 6 Q phase architectures have been identified in small molecule
    amphiphile systems.~\cite{mariani_cubic_1988}
    \item Of these, only 4 are consistent with diffraction data generated
    by the gemini surfactant used here.~\cite{pindzola_cross-linked_2003}
    \item Q$^{230}$ (Ia3d), Q$^{224}$ (Pn3m), Q$^{229}$ (Im3m) or Q$^{212}$ (P4\textsubscript{3}32) phase of type I configuration
    \item The presence of $1 / \sqrt{6}$ and $1 / \sqrt{8}$ peaks rules out the
    Q$^{227}$ (Fd3m) and Q$^{223}$ (Pm3n) configurations.
    \item The Pn3m and Ia3d architectures are the most common~\cite{mariani_cubic_1988,wiesenauer_nanoporous_2012}
    \item The most likely phase is the type I Q$^{230}$ (Ia3d) because it is the
    most common phase observed between lamellar and hexagonal phases and the
    monomeric alkyltrimethylammonium salts used to synthesize the gemini LCs
    exhibit clear Ia3d symmetry.
  \end{itemize}

@@

  There are few molecular simulation studies which study structure and transport
  in LLC systems.
  \begin{itemize}
    \item Mondal et al. used MD simulations to study the self assembly of gemini 
    surfactants and observed the formation of H\textsubscript{II} and Q\textsubscript{I}
    phases depending on water content.\cite{mondal_self-assembly_2013}
    \item Mantha and Yethiraj as well as Roy et al. studied the dynamics of confined 
    water in these systems and showed orders of magnitude differences in their motion
    dependent on system geometry.~\cite{mantha_dynamics_2016,roy_water_2016}
    \item Jackson et al. as well as Mantha et al. have combined experiment with simulation
    in order to show how the choice of monomer head group counterion regulate water 
    dynamics.~\cite{jackson_ion-specific_2018,mantha_counterion-regulated_2018}
    \item Sakamoto et al. and Nada et al. observed the dynamics of water molecules
    and ions and hypothesized transport mechanisms in a simplified model of an LLC
    nanopore.~\cite{sakamoto_development_2018,nada_transport_2020}
  \end{itemize}
